---
title: "NHC Article EDA"
author: Sean Browning (sbrowning <a> cdc <punto> gov)
date: Mar 30, 2022
output: 
    html_document:
        df_print: tibble
        code_folding: hide
---

```{r lib, include = FALSE}
library(devtools)
library(dplyr)
library(readr)
library(stringr)
```

```{r read in}
# Read in latest data pull
raw_data <- read_csv("latest_nhc_data.csv")

print(raw_data)
```

```{r funcs}
# WIP Functions I'll probably include in the package
# once I'm done
para_preclean <- function(x) {
  # # Remove normal line returns
  # out <- gsub("\n", "", x)
  # the apostrophe in Xi'an and others that throws weird errors
  out <- gsub("\u0092", "'", x)

  return(out)
}

para_split <- function(x) {
  # Split on carriage returns
  splits <- strsplit(x, "\n")
  # Remove empty sets
  out <- lapply(splits, function(x) x[x != ""])

  return(out)
}
```

## Investigating "Archetypes"

I'll try just splitting each article by carriage return.

Then, I take the oldest and latest example of each "article type" by paragraph count
where n_content shows how many bulletins had the same number of paragraphs (n_para)
So we can see if there was an iterative change in design over time or if they're
interspersed.

Ideally, what we see is a linear (or at least consistent) usage of a particular type over time

```{r paragraph investigation}
para_summary_df <- raw_data %>%
  mutate(
    content = para_preclean(content),
    para_splits = para_split(content),
    n_para = vapply(para_splits, length, integer(1))
  ) %>%
  group_by(n_para) %>%
  mutate(n_content = n()) %>%
  arrange(desc(date)) %>%
  filter(date %in% range(date)) %>%
  ungroup()

print(para_summary_df)
```

Not exactly as expected, but there is a pattern, and the vast majority can be split into 5 paragraphs.

That style of bulletin apparently first began in 2/14/2020, so that's positive.
We could set an initial value for each region and just write a parser for those bulletins and ignore the rest for the most part.

## Writing a parser for the primary

Let's take a random bulletin and get the data into tabular shape.

```{r select a bulletin to parse}
to_parse <- raw_data %>%
  mutate(
    content = para_preclean(content),
    para_splits = para_split(content),
    n_para = vapply(para_splits, length, integer(1))
  ) %>%
  filter(n_para == 5) %>%
  arrange(desc(date)) %>%
  slice(1)

cat(to_parse$content)
```

The easiest thing to do here is probably to latch onto "in" and then use the right side as the name,
and the left side numbers as the value.

```{r first pass regex}
first_para <- to_parse$para_splits[[1]][1]

str_match_all(first_para, "(([\\d,]*) in ([^\\d,]*) ([,\\)]*|and)*)")
```

